---
title: "Reproducible research - Assignment 1"
author: "ap8980"
date: "24/05/2021"
output: html_document
---

## Loading and preprocessing the data
#import data
#create new data set with removed NA values
#turn date from char to date class
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
activity <- read.csv("activity.csv")
activity_one <- activity %>% drop_na()
activity_one <- activity_one %>% mutate(date = ymd(date))

```

### What is mean total number of steps taken per day?
#calculate mean daily steps

```{r, echo = TRUE}
mean_steps <- activity_one %>% 
group_by(date) %>%
summarise(steps = sum(steps))

mean_steps


```

### What is the average daily activity pattern?
```{r , echo=TRUE}
mean_hist_steps <- ggplot(data = mean_steps)+
    geom_histogram(mapping = aes( x = steps))+
    scale_x_continuous(breaks = seq(0,30000, by = 2000))+
    scale_y_continuous(breaks = seq(0,10, by = 1))+
    ggtitle("Steps per day")

mean_hist_steps

```
```{r, echo = TRUE}
mean(mean_steps$steps)

median(mean_steps$steps)

#average number of steps by intervals
#create new table by intervals
intervals <- activity_one %>%
group_by(interval) %>%
summarise(steps = mean(steps))

```
```{r, echo = TRUE}
#create line chart
ggplot(data = intervals)+
    geom_line(mapping = aes(x = interval, y = steps))+
    scale_x_continuous(breaks = seq(0,2500, by = 200))

```

#interval with max number of steps
```{r, echo = TRUE}
#interval with max number of steps
intervals %>% arrange(desc(steps)) %>% slice_head()
```
### Imputing missing values
```{r}
#sum total NA rows
sum(is.na(activity))

#fill in missing values (use mean)
activity_complete <- activity
nas <- is.na(activity_complete$steps)
avg_interval <- tapply(activity_complete$steps, activity_complete$interval, mean, na.rm=TRUE, simplify=TRUE)
activity_complete$steps[nas] <- avg_interval[as.character(activity_complete$interval[nas])]

#activity complete - new dataset with mean values imputed
steps_complete <- activity_complete %>% 
group_by(date) %>%
summarise(steps = sum(steps))
```
```{r}
#new histogram
ggplot(data = steps_complete)+
    geom_histogram(mapping = aes( x = steps))+
   scale_x_continuous(breaks = seq(0,30000, by = 2000))+
  scale_y_continuous(breaks = seq(0,10, by = 1))+
 ggtitle("Steps per day with imputed values")

```
```{r, echo = TRUE}
#new mean
mean(steps_complete$steps)

#new median
median(steps_complete$steps)
```
### Are there differences in activity patterns between weekdays and weekends?
```{r}
#weekday activity
activity_final <- activity_complete %>% mutate(wday(activity_complete$date, label = TRUE))
#rename column
activity_final<- activity_final %>% rename(weekday = `wday(activity_complete$date, label = TRUE)`)

#summarise pre plot
interval_final <- activity_final %>%
group_by(interval,weekday) %>%
summarise(steps = mean(steps))
```
```{r}
#ggplot via day of week
ggplot(interval_final, aes(x=interval, y=steps, color = weekday)) +
geom_line() +
facet_wrap(~weekday)+
  ggtitle("Activity comparison weekend vs weekdays")
```


